---
title: "Forecasting Time Series"
author: "Daniel Jeffry"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

This reports provides material forecasting demand using Arima, Exponential Smoothing, Croston, and Syntetos Boyland Approximation method. The dataset using in this report for modeling is real case data in one of the state owned company in Indonesia.  

The report as structured as follows:  
1. Data Extraction  
2. Exploratory Data Analysis  
3. Data Preparation  
4. Modeling  
5. Evaluation  
6. Recommendation  

# 1. Data Extraction

Import necessary libraries.

```{r}
library(readxl)
library(tidyverse)
library(fpp3)
library(dplyr)
```

Library **readlxl**: for import excel files into R.  
Library **tidyverse**: for select unique/distinct rows from data frame.      
Library **fpp3**: for visualization of correlation coefficient matrix  
Library **gridExtra**: for plotting multiple graphs. 
Library **caret**: for One Hod Encoding
